### 2019-01-11 paper.tex
empty paper

% no longer than 5 pages including figures and tables + as many pages as needed for references
% 2 columns on 8.5 x 11 inch paper
% 10 point Times Roman font on 12 point single spaced leading, 1 inch margins
% 0.25 inch gutter (seeparation between columns)

% Title, author names, affiliations, abstract should appear on first page
% pages should be numbered
% figures and tables should not require magnification, may contain color but should be
% legible in black and white

\documentclass[10pt,sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\begin{document}

\end{document}

### 2019-01-14 paper.tex
partial draft

@@
 
-\documentclass[10pt,sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[twocolumn,acmtog,10pt,review]{acmart}
+
 \begin{document}
+%\author{}
+%\affiliation{}
+%\author{}
+%\affiliation{}
+
+\section{Introduction}
+
+\section{Analysis tool}
+\label{sec:analysis}
+In this section we describe the tools we use to analyse, diagnose problems, and suggest solutions
+for project builds that use make.  In section \ref{sec:makeanalysis} we will discuss how we use
+a modified version of GNU Make to record enough information about a project's build to later
+reconstruct the build graph during a post mortem analysis.  In section \ref{sec:straceinstr} we
+will discuss how we additionally use a slightly modified version of strace to record all of the
+system calls made during the build and then use this information to determine the real dependencies
+of the build by looking at which files were actually touched. In section \ref{sec:revisegraphs} we
+will look at how we combine the information from these two tools to transform build graphs into
+new graphs with potentially more parallelism and provide insight into the build graph's structure.
+% what else to say here?
+
+\subsection{Make analysis and instrumentation}
+\label{sec:makeanalysis}
+
+To collect the desired level of information from a run of GNU Make \cite{}, we run make with the
+debug level of 'verbose', use our own make wrappers which print additional information, and use
+a modified version of make which prints directory and timing information for each shell command
+it runs.  All of this information is written to a file and is used by our tool to reconstruct
+the build graph after the build successfully completes.  We are able to re-construct a graph which
+contains all of the shell commands executed to build the project as well as the time it took
+to execute each one.  With this timing information we can calculate the work and the span of the
+build.  The work is the total time to execute every shell command in the build graph.  The span is
+the longest time to execute any sequence of shell commands in the build graph which must be
+executed sequentially.  For example: the span of a make rule is the longest time to execute a
+dependency plus the sum of the times to execute each recipe.  And parallelism is defined to be
+span/work, therefore to increase the parallelism of a project and reduce the overall running time,
+the span must be decreased.
+
+Identifying the span path in the build allows the user to target the most appropriate place
+for performance improvements.  In section \ref{} we will discuss some projects that were analysed
+using this tool and what we learned about their parallelism and how to improve it.
+
+% Have a wrapper around gnu make; which prints information about which top-level make
+% is being invocated and prints out the directory; this wrapper also times the call to
+% actual gnu make
+
+% have a 2nd wrapper around gnu make; which is called for ``sub makes''. we tell actual
+% make that we want it to call this for all $(MAKE) calls in the makefile.  this wrapper
+% similarly to the top-make wrapper prints out which make is being executed as well as the
+% directory and time of the actual make call.
+% WHY do we need the above information??
+
+% The most important change is in make itself; where we modified make to print each shell
+% command it is running as well as the time the shell command took.
+% this data lets us calculate the work and the span of the build and lets us attribute span
+% to the correct targets and shell commands.
+% Since we are able to store the command run by make in the reconstructed build graph we can
+% rerun those ocmmands to replay the build; and it didn't require us to parse the makefile
+% ourselves.
+
+% tool can calculate work and span of the build and attribute the span to the correct nodes
+% so that the user can focus their performance improvements on the portiosn of the graph that
+% will have the most affect.
+
+\subsection{Strace instrumentation}
+\label{sec:straceinstr}
+
+In order to extend the analysis we can perform on build graphs, we use strace \cite{} in
+conjunction with our enhanced version of GNU Make.  Strace lets us capture and record the system
+calls of every process launched during a make build so we can later map directly from shell
+commands launched by make to the system calls those shell commands ran.  Strace can write the
+system calls and signals from each pid to a different file with the pid as the file extension.
+One problem with this is that Linux may re-use pids during the course of a build, which by default
+causes strace to overwrite the existing file.  To solve this problem and maintain a 1-to-1
+mapping between strace data and processes launched by make, we modified strace to append a
+timestamp to each file extension along with the pid.
+
+Once the correct system call data is mapped to the correct build command, the strace data is
+processed to determine which files were inputs and outputs of that command.  An input in this case
+is a file that could have been read but not written and an output is a file that could have been
+written.  Using this information we can modify the build graph and we will discuss this more
+in the next section.
+
+% strace lets us capture and record the system calls of every process launched during a make build
+% and it will save each process pid's system calls to an individual file which makes it easy for
+% us to map each shell command to its system calls.  There is one wrinkle in this plan though,
+% pids can be reused and pid reuse is not strictly increasing and the version of strace that we
+% used would overwrite a file if this occured which results in inaccurate data.
+
+% to solve this problem we have strace also append a timestamp to the end of the output file
+% extensions.  This results in unique output files which are also strictly increasing with
+% respect to the timestamps.
+
+% we then parse the strace information to determine which files were touched by a process and
+% we classify these files as input files and as output files.  Two processes which only share
+% input files should in theory be able to run in parallel.  So using this information we can
+% modify the make graph to allow those two processes to run in parallel.
+
+\subsection{Revising graphs}
+\label{sec:revisegraphs}
+
+The build graph is made up of two types of vertices and two types of directed edges.  A node is
+either a leaf, which represents a file or a single shell command executed by make, or a non-leaf,
+which represents a single target.  An edge is either a dependency edge, which indicates that the
+vertice the edge goes to can run in parallel with other vertices reach by dependency edges from the
+same source vertice, or a recipe edge, which indicates that the vertice the edge points to may not
+run in parallel with other vertices reachable from the source vertice, and it must run in a
+specific order relative to the other vertices reachable from the source.
+
+To increase the span of the build graph we would like to transform recipe edges into dependency
+edges if possible.  To determine if this is possible we use the collected system call data and
+calculate the file inputs and file outputs of each shell command.  If two adjacent shell commands
+which occur sequentially, share zero or more input files and share no input and output files, then
+those two commands can run in parallel.
+
+Our graph re-builder performs this analysis on the original build graph and produces a new build
+graph which hopefully has a lower span than the original graph.
+
+% build graph is made up of two types of node and two types of edges.  A node is either a
+% leaf node which represents a file or represents a single shell command executed by make, and
+% a non-leaf node which represents a single target.  And an edge is either a dependency edge or
+% a recipe edge.  A dependency edge points to another node which must be run before recipe
+% edges of the node may be run.  All dependnecy edges may be run in parallel while the recipe
+% edges must be run sequentially, so there is an explicit ordering on recipe edges.
+
+% To increase the parallelism of a build graph we want to transform recipe edges into
+% dependency edges if possible.
+
+% to do this we combine the build graph we constructed with the strace data we collected to
+% compare the input and output files of each recipe run as well as each command.  If two
+% adjacent
+
+\section{Make build analyses}
+
+In this section we will look at four problems with Make build systems.  First we will discuss
+what it means for a build to be unnecessarily sequential and how we can diagnose and suggestion
+solutions to that problem.  Second, we will discuss build systems being dominated by a single
+expensive command that prevents the build from running more quickly and how we can diagnose and
+suggest solutions to that problem.  Third, we will discuss when a build has ample available
+parallelism but doesn't achieve its maximum paralleism due to bad operating system scaling.
+Fourth, we will discuss the problem of a slow compiler.
+
+\subsection{Unnecessarily sequential}
+
+Makefiles are a collection of rules which describe how to build some targets.  Each rule consists
+of a a series of prerequisites and a series of recipes which are run to build the target.  The
+prerequisites can be built in parallel but the recipes are run in the order they are listed.  So,
+the author of a makefile can affect the potential parallelism of the build by how they write their
+makefile.
+
+A makefile which is too sequential will suffer from an excessively large span.
+% not sure what to say here
+
+To diagnose this problem using our tool, the user can calculate the span of the original build
+graph and then let the tool build a new graph using strace data and calculate the span of the
+new graph.  If the span of the new graph is less than the span of the original graph then the
+original graph was unnecessarily sequential.
+
+The way to fix a graph like this is for the makefile writer to re-write the makefile to
+remove the false-dependencies discovered by the tool.
+
+Example: todo
+
+\subsection{Expensive command}
+
+All build systems which lack the desired parallelism can be characterized as having too great of
+a span in relation to their work.  The span in a build system is the longest (in time) series of
+tasks that must be executed sequentially.  In some cases the span is not a series of similar length
+tasks but is dominated by a single long running task.  In cases like these it is best to point the
+user to the most expensive task in the span because reducing the running time of that task is the
+best way to increase the parallelism of the build.
+
+Reducing the running time of non-span tasks will not decrease the running time of the project as
+a whole.
+
+To diagnose this problem the user can use the tool to discover the span path through the build
+graph and see which target and shell commands are the most time intensive.  The user can then see
+what sort of improvement would occur to the span if they were able to decrease the running
+time of any target in the graph.  This is useful in the case where decreasing the running time
+of a single task gives the build graph a new span path which is dominated by a different also
+long running command that prevents the span of the graph from decreasing.
+
+An example of such a project with such a build is Facebook's zstd project \cite{}.  This project
+has more than one long running CC command provide over 90\% of the running time of the build's
+span.  Splitting the CC command into many parallel CC commands followed by a linking command
+and then having the tool build a new graph presented a new span that was also dominated by
+another long running CC command.
+
+\subsection{Bad OS Scaling}
+
+Some projects may exhibit ample potential parallelism, but may be limited by systems that do not
+have the resources to exploit said parallelism.  LLVM \cite{}
+
+% project example is LLVM:
+% say something about how LLVM is a very large project and
+
+\subsection{Slow compiler}
+
+\section{Future work}
+
+One of the limitations of this work is that currenlty to collect the necessary data to rebuild
+the build graph, the project must be built using a single thread.  This is a burden for projects
+that are very long running such as LLVM.  Additionally, building with strace causes an additional
+slowdown and inflates the running time of each shell command, which makes the raw numbers
+reported by a build graph created with strace on, inaccurate.  In the future for performance sake
+it would be good to collect the necessary information while allowing make to run in parallel with
+any number of processes.
+
+In addition to this, we hope to analyse more project builds and introduce parallelism at other
+level of project builds, such as at the level of the compiler.
 

### 2019-01-15 paper.tex
Add title.
abstract.

@@
 
-\documentclass[twocolumn,acmtog,10pt,review]{acmart}
+\documentclass[twocolumn,sigplan,10pt,review]{acmart}
+\title{Four reasons your build is slow}
+\author{Sarah Spall}
+\affiliation{Indiana University}
+\author{Sam Tobin-Hochstadt}
+\affiliation{Indiana University}
+
+\begin{abstract}
+Spending time waiting for code to compile is cliche enough to be
+featured in XKCD, yet it continues to plague software
+developers. Fortunately, software builds are an embarassingly parallel
+problem, and so we can finally make use of all those extra cores---a
+happy coincidence of software complexity and hardware trends.
+
+Unfortunately, the story is not quite this rosy, as almost anyone who
+waits for software to build knows. Real build systems fail to be the
+embarassingly parallel poster child we hope for, and instead are slow
+in a wide variety of ways.
+
+To investigate this problem, we have built an automated build analysis
+tool that works for arbitrary build systems using the venerable
+\texttt{make} tool. Our analysis can predict parallel speedup, analyze
+the long pole of dependencies, determine false dependencies, and
+predict parallelism in a perfect world. We use our tool on real-world
+project to describe four problems found in actual software builds, and
+in some cases how to fix them.
+\end{abstract}
 
 \begin{document}
-%\author{}
-%\affiliation{}
-%\author{}
-%\affiliation{}
+
+\maketitle
 

### 2019-01-15 paper.tex
work on seconds 2 and 3

--- 2019-01-14 paper.tex
@@
 
-\documentclass[twocolumn,acmtog,10pt,review]{acmart}
+\documentclass[twocolumn,sigplan,10pt,review]{acmart}
+\title{Four reasons your build is slow}
+\author{Sarah Spall}
+\affiliation{Indiana University}
+\author{Sam Tobin-Hochstadt}
+\affiliation{Indiana University}
 
 \begin{document}
-%\author{}
-%\affiliation{}
-%\author{}
-%\affiliation{}
+
+\maketitle
 
@@
 
-An example of such a project with such a build is Facebook's zstd project \cite{}.  This project
-has more than one long running CC command provide over 90\% of the running time of the build's
-span.  Splitting the CC command into many parallel CC commands followed by a linking command
-and then having the tool build a new graph presented a new span that was also dominated by
-another long running CC command.
+An example of a project with such a build is Facebook's zstd project \cite{}.  If we use our tool
+to analyse the span of zstd's build, we see that the original span is dominated by a single call
+to cc.
+\begin{verbatim}
+  cc -I../lib -I../lib/common -I../lib/compress -I../lib/dictBuilder -DXXH_NAMESPACE=ZSTD_ -I../lib/legacy -DZSTD_LEGACY_SUPPORT=5 -O3 -DBACKTRACE_ENABLE=0 ../lib/common/debug.c ../lib/common/entropy_common.c ../lib/common/error_private.c ../lib/common/fse_decompress.c ../lib/common/pool.c ../lib/common/threading.c ../lib/common/xxhash.c ../lib/common/zstd_common.c ../lib/compress/fse_compress.c ../lib/compress/hist.c ../lib/compress/huf_compress.c ../lib/compress/zstd_compress.c ../lib/compress/zstd_double_fast.c ../lib/compress/zstd_fast.c ../lib/compress/zstd_lazy.c ../lib/compress/zstd_ldm.c ../lib/compress/zstd_opt.c ../lib/compress/zstdmt_compress.c ../lib/decompress/huf_decompress.c ../lib/decompress/zstd_decompress.c ../lib/dictBuilder/cover.c ../lib/dictBuilder/divsufsort.c ../lib/dictBuilder/fastcover.c ../lib/dictBuilder/zdict.c ../lib/legacy/zstd_v05.c ../lib/legacy/zstd_v06.c ../lib/legacy/zstd_v07.c zstdcli.o fileio.o bench.o datagen.o dibio.o -o zstd
+\end{verbatim}
+
+If this call is no longer part of the span, we find that the new span is similarly dominated by
+another single call to cc.
+
+\begin{verbatim}
+cc -I.
+-I./common -DXXH_NAMESPACE=ZSTD_ -I./legacy -DZSTD_LEGACY_SUPPORT=5 -O3 common/debug.c common/entropy_common.c common/error_private.c common/fse_decompress.c common/pool.c common/threading.c common/xxhash.c common/zstd_common.c compress/fse_compress.c compress/hist.c compress/huf_compress.c compress/zstd_compress.c compress/zstd_double_fast.c compress/zstd_fast.c compress/zstd_lazy.c compress/zstd_ldm.c compress/zstd_opt.c compress/zstdmt_compress.c decompress/huf_decompress.c decompress/zstd_decompress.c deprecated/zbuff_common.c deprecated/zbuff_compress.c deprecated/zbuff_decompress.c dictBuilder/cover.c dictBuilder/divsufsort.c dictBuilder/fastcover.c dictBuilder/zdict.c legacy/zstd_v05.c legacy/zstd_v06.c legacy/zstd_v07.c -shared -fPIC -fvisibility=hidden -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+\end{verbatim}
+
+If their of these calls were members of the build's span, than our tool predicts the span would be
+7.2 seconds rather than 39 seconds or 37 seconds (todo check that time).  To test this we manually
+replaced the original calls with many calls that first compiled each file individually then later
+linked them.  % todo fill in results.
 
@@
 
-Some projects may exhibit ample potential parallelism, but may be limited by systems that do not
-have the resources to exploit said parallelism.  LLVM \cite{}
+Some projects may exhibit ample potential parallelism, but even when run with more threads than
+available parallelism, predicted parallelism may not be achieved.  LLVM \cite{} is the largest
+project we analysed with a calculated work of approximately 12584 seconds, or 3.5 hours.  The
+predicted span of LLVM is 305 seconds, or 5 minutes.  This gives LLVM a parallelism of about 41.
+These calculations where made from a build of LLVM that was done not using strace.
+
+A parallelism of 41 would lead one to believe that on a machine with 32 cores LLVM should be built
+in parallel in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
+built with ``-j 32'' the observed speedup was slightly over 18; see chart \ref{}.  Using work and
+span our tool can predict the potential speedup a build will experience using Brent's theorem
+\cite{}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
+P processors, \begin{math} T_P \leq T \end{math}
+
+% say something about brent's theorem
+% say something about why llvm doesnt achieve the predicted speedup
 

### 2019-01-15 paper.tex

@@
 \affiliation{Indiana University}
+
+\begin{abstract}
+Spending time waiting for code to compile is cliche enough to be
+featured in XKCD, yet it continues to plague software
+developers. Fortunately, software builds are an embarassingly parallel
+problem, and so we can finally make use of all those extra cores---a
+happy coincidence of software complexity and hardware trends.
+
+Unfortunately, the story is not quite this rosy, as almost anyone who
+waits for software to build knows. Real build systems fail to be the
+embarassingly parallel poster child we hope for, and instead are slow
+in a wide variety of ways.
+
+To investigate this problem, we have built an automated build analysis
+tool that works for arbitrary build systems using the venerable
+\texttt{make} tool. Our analysis can predict parallel speedup, analyze
+the long pole of dependencies, determine false dependencies, and
+predict parallelism in a perfect world. We use our tool on real-world
+project to describe four problems found in actual software builds, and
+in some cases how to fix them.
+\end{abstract}
 

### 2019-01-15 paper.tex
Demo verb, add some stuff.

--- 2019-01-15 paper.tex
@@
 
-\documentclass[twocolumn,sigplan,10pt,review]{acmart}
+\documentclass[sigplan,10pt,review,authorversion]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
+\citestyle{acmnumeric}
+
+% REMOVE FOR ACCEPTED
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+%\pagestyle{plain}
+
+\expandafter\let\csname not=\endcsname\relax
+\expandafter\let\csname not<\endcsname\relax
+\expandafter\let\csname not>\endcsname\relax
+
+\RequirePackage{pdf14}
+\usepackage{unicode-math}
+
+\usepackage{balance}
+%% \usepackage{fixltx2e} % travis-ci has an old version of latex and needs this for \textsubscript, etc
+\usepackage{mdframed}
+%% \usepackage{booktabs}
+%% \usepackage{enumitem}
+%% \usepackage{amsmath}
+%% \usepackage{amsthm}
+%% \usepackage{color}
+%% \usepackage{calc}
+\usepackage{xspace}
+%% \usepackage{upgreek}
+\usepackage{graphicx}
+%% \usepackage{natbib}
+%% \usepackage{alltt}
+\usepackage{fancyvrb}
+%\usepackage{array}
+%% \usepackage{tabu}
+%% \usepackage{multirow}
+\usepackage{mathpartir}
+%% \usepackage{balance}
+\usepackage{subcaption}
+\usepackage{wrapfig}
+%% \usepackage{framed}
+\usepackage{colortbl}
+\usepackage{tikz}
+\usetikzlibrary{decorations.markings}
+\usetikzlibrary{arrows,shapes}
+\usetikzlibrary{positioning}
+\tikzset{
+    >=stealth,
+    %auto,
+    %node distance=3.5cm,
+    font=\scriptsize,
+    ack/.style={align=center,ellipse,draw},%circle,draw,thick,align=center},
+    minimum size=20pt
+}
+\usepackage{multicol}
+\usepackage{pgfplots}\pgfplotsset{compat=1.13}
+
+%% Requires:
+%% \usepackage{color}
+%% \usepackage{calc}
+%% \usepackage{xspace}
+\usepackage{pifont}
+
+\usepackage{tikz-cd}
+
+%% \usepackage{upgreek}
+
+% \usepackage{MnSymbol}
+
+\setlength{\bibsep}{1pt}
+
+%\newtheorem{theorem}{Theorem}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{definition}{Definition}
+
+\definecolor{gray}{rgb}{0.9,0.9,0.9}
+\definecolor{red}{rgb}{1,0,0}
+\newcommand{\graybox}[1]{\mbox{\setlength{\fboxsep}{0.5pt}%
+    \colorbox{gray}{$#1$}}}
+
+%\newcommand{\graybox}[1]{#1}
+
+\newcommand{\var}[1]{\mathit{#1}}
+
+\newcommand{\fixme}[1][\relax]{{\color{red}{FIXME: #1}}}
+
+\newcommand{\ma}[1]{\ensuremath{#1}\xspace}
+
 \title{Four reasons your build is slow}
@@
 \section{Introduction}
+
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+all:
+  cc in.c -o out.o
+\end{Verbatim}
 

### 2019-01-16 paper.tex
adding figures and working on zstd section

@@
 
+\begin{figure}[t]
+  \includegraphics[width=0.5\textwidth]{zstd-speedup}
+  \caption{todo}
+  \label{fig:zstd}
+\end{figure}
+
 All build systems which lack the desired parallelism can be characterized as having too great of
@@
 to cc.
-\begin{verbatim}
-  cc -I../lib -I../lib/common -I../lib/compress -I../lib/dictBuilder -DXXH_NAMESPACE=ZSTD_ -I../lib/legacy -DZSTD_LEGACY_SUPPORT=5 -O3 -DBACKTRACE_ENABLE=0 ../lib/common/debug.c ../lib/common/entropy_common.c ../lib/common/error_private.c ../lib/common/fse_decompress.c ../lib/common/pool.c ../lib/common/threading.c ../lib/common/xxhash.c ../lib/common/zstd_common.c ../lib/compress/fse_compress.c ../lib/compress/hist.c ../lib/compress/huf_compress.c ../lib/compress/zstd_compress.c ../lib/compress/zstd_double_fast.c ../lib/compress/zstd_fast.c ../lib/compress/zstd_lazy.c ../lib/compress/zstd_ldm.c ../lib/compress/zstd_opt.c ../lib/compress/zstdmt_compress.c ../lib/decompress/huf_decompress.c ../lib/decompress/zstd_decompress.c ../lib/dictBuilder/cover.c ../lib/dictBuilder/divsufsort.c ../lib/dictBuilder/fastcover.c ../lib/dictBuilder/zdict.c ../lib/legacy/zstd_v05.c ../lib/legacy/zstd_v06.c ../lib/legacy/zstd_v07.c zstdcli.o fileio.o bench.o datagen.o dibio.o -o zstd
-\end{verbatim}
+
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  cc -I../lib -I../lib/common -I../lib/compress -I../lib/dictBuilder 
+     -DXXH\_NAMESPACE=ZSTD\_ -I../lib/legacy -DZSTD\_LEGACY\_SUPPORT=5 -O3 
+     -DBACKTRACE\_ENABLE=0 
+     ../lib/common/debug.c ../lib/common/entropy\_common.c ../lib/common/error\_private.c 
+     ../lib/common/fse\_decompress.c ../lib/common/pool.c ../lib/common/threading.c 
+     ../lib/common/xxhash.c ../lib/common/zstd\_common.c ../lib/compress/fse\_compress.c 
+     ../lib/compress/hist.c ../lib/compress/huf\_compress.c 
+     ../lib/compress/zstd\_compress.c ../lib/compress/zstd\_double\_fast.c 
+     ../lib/compress/zstd\_fast.c ../lib/compress/zstd\_lazy.c 
+     ../lib/compress/zstd\_ldm.c ../lib/compress/zstd\_opt.c 
+     ../lib/compress/zstdmt\_compress.c ../lib/decompress/huf\_decompress.c 
+     ../lib/decompress/zstd\_decompress.c ../lib/dictBuilder/cover.c 
+     ../lib/dictBuilder/divsufsort.c ../lib/dictBuilder/fastcover.c 
+     ../lib/dictBuilder/zdict.c ../lib/legacy/zstd\_v05.c ../lib/legacy/zstd\_v06.c 
+     ../lib/legacy/zstd\_v07.c zstdcli.o fileio.o bench.o datagen.o dibio.o -o zstd
+\end{Verbatim}
 
@@
 
-\begin{verbatim}
+\begin{verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
 cc -I.
@@
 
-If their of these calls were members of the build's span, than our tool predicts the span would be
-7.2 seconds rather than 39 seconds or 37 seconds (todo check that time).  To test this we manually
-replaced the original calls with many calls that first compiled each file individually then later
-linked them.  % todo fill in results.
+If neither of these calls were members of the build's span, then our tool predicts the span would be
+7.2 seconds rather than 39 seconds or 37 seconds.  To test this we manually replaced the first of
+these two cc calls with 28 new calls to cc; 27 of those are spent compiling each .c file individually
+and the last is to link them all together.  Due to this new organization the 27 .c files can be
+compiled in parallel.  Similarly, the second cc call is replace with 28 new calls to cc; 27 of which
+compile each .c file individually and the last of which links the object files together.
+
+You can see in figure \ref{fig:zstd} that when we ran the modified build with 1 to 32 jobs we
+observed the predicted speedup.  An additional performance improvement that we did not predict was
+that the work of the new build would decrease from approximately 120 seconds to approximately 80 seconds.
+
+Each of the generated object files is declared as its own target; so when make considers building
+each object file it can check if the file already exists and whether it must be rebuilt.  It turns
+out that the object files being built as part of the first long cc command, had already been
+previously built by make.  By declaring each object file as its own target and building each one
+individually then linking them, make was able to determine it wasn't necessary to build them the
+2nd time; so the effect time of the first cc command when from 40 seconds to 2 seconds; only the
+time necessary for linking the object files together.
 
 \subsection{Bad OS Scaling}
+
+\begin{figure}[]
+  \includegraphics[width=0.5\textwidth]{llvm-speedup}
+  \caption{todo}
+  \label{fig:llvm}
+\end{figure}
 

### 2019-01-16 paper.tex
work on slow compiler section

@@
 \subsection{Expensive command}
+\label{sec:expensive}
 
@@
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  cc -I../lib -I../lib/common -I../lib/compress -I../lib/dictBuilder 
-     -DXXH\_NAMESPACE=ZSTD\_ -I../lib/legacy -DZSTD\_LEGACY\_SUPPORT=5 -O3 
-     -DBACKTRACE\_ENABLE=0 
-     ../lib/common/debug.c ../lib/common/entropy\_common.c ../lib/common/error\_private.c 
-     ../lib/common/fse\_decompress.c ../lib/common/pool.c ../lib/common/threading.c 
-     ../lib/common/xxhash.c ../lib/common/zstd\_common.c ../lib/compress/fse\_compress.c 
-     ../lib/compress/hist.c ../lib/compress/huf\_compress.c 
-     ../lib/compress/zstd\_compress.c ../lib/compress/zstd\_double\_fast.c 
-     ../lib/compress/zstd\_fast.c ../lib/compress/zstd\_lazy.c 
-     ../lib/compress/zstd\_ldm.c ../lib/compress/zstd\_opt.c 
-     ../lib/compress/zstdmt\_compress.c ../lib/decompress/huf\_decompress.c 
-     ../lib/decompress/zstd\_decompress.c ../lib/dictBuilder/cover.c 
-     ../lib/dictBuilder/divsufsort.c ../lib/dictBuilder/fastcover.c 
-     ../lib/dictBuilder/zdict.c ../lib/legacy/zstd\_v05.c ../lib/legacy/zstd\_v06.c 
-     ../lib/legacy/zstd\_v07.c zstdcli.o fileio.o bench.o datagen.o dibio.o -o zstd
+  cc \$(FLAGS)
+  ../lib/common/debug.c
+  ../lib/common/entropy\_common.c
+  \textrm{\emph{25 more .c files}}
+  zstdcli.o fileio.o bench.o
+  datagen.o dibio.o -o zstd
 \end{Verbatim}
@@
 
-\begin{verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-cc -I.
--I./common -DXXH_NAMESPACE=ZSTD_ -I./legacy -DZSTD_LEGACY_SUPPORT=5 -O3 common/debug.c common/entropy_common.c common/error_private.c common/fse_decompress.c common/pool.c common/threading.c common/xxhash.c common/zstd_common.c compress/fse_compress.c compress/hist.c compress/huf_compress.c compress/zstd_compress.c compress/zstd_double_fast.c compress/zstd_fast.c compress/zstd_lazy.c compress/zstd_ldm.c compress/zstd_opt.c compress/zstdmt_compress.c decompress/huf_decompress.c decompress/zstd_decompress.c deprecated/zbuff_common.c deprecated/zbuff_compress.c deprecated/zbuff_decompress.c dictBuilder/cover.c dictBuilder/divsufsort.c dictBuilder/fastcover.c dictBuilder/zdict.c legacy/zstd_v05.c legacy/zstd_v06.c legacy/zstd_v07.c -shared -fPIC -fvisibility=hidden -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
-\end{verbatim}
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  cc \$(FLAGS)
+  common/debug.c
+  common/entropy\_common.c
+  \textrm{\emph{28 more .c files}}
+  -shared -fPIC -fvisibility=hidden
+  -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+\end{Verbatim}
 
@@
 
+A similar issue to the one described in section \ref{sec:expensive} is when a build's
+span is dominated by the compilation of a single file.  This problem however cannot be
+solved in the makefile like the zstd issue was.  Solutions to this problem include
+breaking the file up into smaller files which can be compiled in parallel and then
+linked; or changing the compiler so that it is capable of compiling the file faster.
+A faster compiler, would have the additional benefit of not only decreasing the span
+of the project but also likely decreasing the work of the project.
+
+Google's protocol buffers project \cite{} exhibited this behavior.  It had a relatively
+long predicted span of ?? seconds, which was dominated by a call to g++ which was
+compiling a single file.  When we apply the same analysis as we applied to zstd in
+section \ref{sec:expensive}, we see that if \ref{} wasn't a part of the span, the new
+span would be ?? and would be dominated by another g++ compilation of a single file.
+And we continue this again to find that the 3 top span paths are dominated by g++
+compilations of individual files.  If none of these calls were part of the span, the
+new span of the project would be ??.  Now if the developers of protocol buffers were
+interested in increasing the parallelism of their project they would know where they
+should focus their efforts
+
+Here is a simplification of the real command run by make; it took about 40 seconds to run.
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  g++ \$(FLAGS)
+  -o google/protobuf/descriptor.lo
+  google/protobuf/descriptor.cc
+\end{Verbatim}
+
+Here is a simplification of the real command run by make; it took about 27 seconds to run.
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  g++ \$(FLAGS)
+  -o google/protobuf/descriptor.pb.lo
+  google/protobuf/descriptor.pb.cc
+\end{Verbatim}
+
+Here is a simplification of the real command run by make; it took about 22.5 seconds to run.
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  g++ \$(FLAGS)
+  -o google/protobuf/compiler/cpp/cpp\_message.lo
+  google/protobuf/compiler/cpp/cpp\_message.cc
+\end{Verbatim}
+
+% so, if the problem is with compiling a single file, then we don't really have a
+% make level solution to that problem
+% user needs to do some work at the file/compiler level to create a solution
+% potential solutions include breaking the file up into smaller files which can be
+% compiled and linked in paralle
+% using a faster compiler, which would have the added benefit of likely decreasing
+% the overall work of the build as well as the span
+% maybe trying to motivate some future work here
+
+% example project is protobufs c++ version;
+% We perform the same analysis that we did in section expensive;
+% look at the top spans and what they are dominated by
+% we see that the top 3? are dominated by g++ compilations of single files
+% and if those are no longer part of the span we half our span; which would result in a
+% double speedup
+% so we now have a place to focus our energy for improvements
+% HIgh level; non make level fix; expensive section is a make level fix; should mention that
+% there maybe
+
 \section{Future work}

### 2019-01-17 paper.tex
work on section 'unnecessarily sequential'

@@
 
-\subsection{Unnecessarily sequential}
-
-Makefiles are a collection of rules which describe how to build some targets.  Each rule consists
-of a a series of prerequisites and a series of recipes which are run to build the target.  The
-prerequisites can be built in parallel but the recipes are run in the order they are listed.  So,
-the author of a makefile can affect the potential parallelism of the build by how they write their
-makefile.
-
-A makefile which is too sequential will suffer from an excessively large span.
-% not sure what to say here
-
-To diagnose this problem using our tool, the user can calculate the span of the original build
-graph and then let the tool build a new graph using strace data and calculate the span of the
-new graph.  If the span of the new graph is less than the span of the original graph then the
-original graph was unnecessarily sequential.
-
-The way to fix a graph like this is for the makefile writer to re-write the makefile to
-remove the false-dependencies discovered by the tool.
-
-Example: todo
-
 \subsection{Expensive command}
@@
 time necessary for linking the object files together.
+
+\subsection{Unnecessarily sequential}
+
+Makefiles are a collection of rules which describe how to build some targets.  Each rule consists
+of a a series of prerequisites and a series of recipes which are run to build the target.  The
+prerequisites can be built in parallel but the recipes are run in the order they are listed.  So,
+the author of a makefile can affect the potential parallelism of the build by how they write their
+makefile.
+
+A makefile which is too sequential will have a longer than necessary span.  In the previous
+section we talked about projects having a long span because it was dominated by a single
+long running shell command.  Another possiblility is the span being made up of many shell
+commands; none of which dominate.  In the previous section we talked about zstd, and if we
+rewrite the makefiles so the long running commands are made up of many sequential compilation
+and then linking commands we end up with a build that has the same work and span as the original.
+
+To diagnose this problem using our tool, the user can once again calculate the span of the build
+graph and look at what targets and shell calls constitute the span.  In this case the span is
+made up of many short running compilation jobs.  To check if these compilation jobs may run in
+parallel, our tool will take the original graph, the collected strace data, and build a new graph
+as described in section \ref{sec:revisegraphs}.  The span of the new graph can then be compared
+to the span of the old graph; if the span of the new graph is less than the span of the old graph,
+then the original graph was unnecessarily sequential.  The allow this build to run faster in the
+future, the developer can rewrite the makefile to reflect the organization of the build in
+the new graph.
+
+In the case of zstd, when built with strace; the calculated work for the original graph is 133
+seconds and the span is 45 seconds; which means a parallelism of 2.95.  When this graph is
+rebuilt using strace data and the span is recalculated, it is 8 seconds, which matches with the
+predictions and observations from section \ref{sec:expensive}.
 

### 2019-01-17 paper.tex
Intro.

@@
 
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-all:
-  cc in.c -o out.o
-\end{Verbatim}
+If you develop software, or even if you compile it, you probably run
+into software that's slow to build. This problem is such a cliche that
+it appears in comics, but it's also such a problem that companies
+routinely outfit developers with the beefiest machines possible---all
+to compile software projects faster. Fortunately, for most systems and
+languages, compiling software is a classic example of embarassing
+parallelism. As a result, developers who spend much of their time
+compiling large software projects use highly parallel systems to do so
+[cite Regehr tweet].
+
+This is perhaps best demonstrated by the classic \verb|make -j|
+command, which builds on \verb|make|~\cite{make} to execute
+dependencies in parallel, consuming as many parallel resources as are
+available. The process-oriented nature of \verb|make| and the graph
+structure of dependencies combine for easy and effective parallelism.
+
+However, this seeming solution rarely works out as well in practice as
+it does in theory. In addition to the substantial expense of
+highly-parallel systems, real software rarely makes use of the full
+capacity of the machine it is built on, despite the potential for high
+degrees of parallelism.
 

### 2019-01-17 paper.tex
round 1 of editing

@@
 \label{sec:analysis}
-In this section we describe the tools we use to analyse, diagnose problems, and suggest solutions
-for project builds that use make.  In section \ref{sec:makeanalysis} we will discuss how we use
-a modified version of GNU Make to record enough information about a project's build to later
-reconstruct the build graph during a post mortem analysis.  In section \ref{sec:straceinstr} we
-will discuss how we additionally use a slightly modified version of strace to record all of the
-system calls made during the build and then use this information to determine the real dependencies
-of the build by looking at which files were actually touched. In section \ref{sec:revisegraphs} we
-will look at how we combine the information from these two tools to transform build graphs into
-new graphs with potentially more parallelism and provide insight into the build graph's structure.
-% what else to say here?
+In this section we describe the tools we use to analyse make builds, diagnose problems related
+to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
+discuss the modified version of GNU Make \cite{} we use to record enough information about a
+project's build to reconstruct the build graph during a post-mortem analysis.
+In section \ref{sec:straceinstr} we discuss how we use a slightly modified version of strace to
+record all of the system calls made during the project's build and then use the system calls
+to determine the real dependencies of the build. In section \ref{sec:revisegraphs} we discuss
+how to combine the information from these two tools to transform a build graph into
+a new graph with potentially more parallelism.
 
@@
 
-To collect the desired level of information from a run of GNU Make \cite{}, we run make with the
-debug level of 'verbose', use our own make wrappers which print additional information, and use
-a modified version of make which prints directory and timing information for each shell command
-it runs.  All of this information is written to a file and is used by our tool to reconstruct
-the build graph after the build successfully completes.  We are able to re-construct a graph which
-contains all of the shell commands executed to build the project as well as the time it took
-to execute each one.  With this timing information we can calculate the work and the span of the
-build.  The work is the total time to execute every shell command in the build graph.  The span is
-the longest time to execute any sequence of shell commands in the build graph which must be
-executed sequentially.  For example: the span of a make rule is the longest time to execute a
-dependency plus the sum of the times to execute each recipe.  And parallelism is defined to be
-span/work, therefore to increase the parallelism of a project and reduce the overall running time,
-the span must be decreased.
+To collect the desired level of information from a run of GNU Make \cite{}, we run it with the
+debug level of 'verbose', use our own wrappers, which print additional information, and use
+a modified version of make which prints directory and timing information for each shell command run.
+All of this information is written to a file and later used to reconstruct
+the build graph of a successful build.  The reconstructed graph contains all of the shell commands
+executed during the build as well as the time each one took to complete.  With the structure of the
+build and the timing information the work and span of the build can be calculated. Work is the total
+time to execute every shell command in the build graph, and span is the longest time to execute any
+sequence of shell commands which must execute sequentially.  For example, the span of a single
+make rule is, the longest time to execute a dependency plus the sum of the times to execute each
+recipe.  Parallelism is defined as \begin{math} \frac{span}{work} \end{math}, therefore to increase
+the parallelism of a project the span must be decreased; making span an important place to focus
+performance improvements.
 
-Identifying the span path in the build allows the user to target the most appropriate place
-for performance improvements.  In section \ref{} we will discuss some projects that were analysed
-using this tool and what we learned about their parallelism and how to improve it.
+Identifying the span path in a build identifies the most important places for performance
+improvements.  In section \ref{sec:realanalyses} we discuss four causes of parallelism
+performance issues in make based project builds, how to use this tool to identify the
+issue and use information about the span of the build to implement a solution.
 
@@
 
-In order to extend the analysis we can perform on build graphs, we use strace \cite{} in
-conjunction with our enhanced version of GNU Make.  Strace lets us capture and record the system
-calls of every process launched during a make build so we can later map directly from shell
-commands launched by make to the system calls those shell commands ran.  Strace can write the
-system calls and signals from each pid to a different file with the pid as the file extension.
-One problem with this is that Linux may re-use pids during the course of a build, which by default
-causes strace to overwrite the existing file.  To solve this problem and maintain a 1-to-1
-mapping between strace data and processes launched by make, we modified strace to append a
-timestamp to each file extension along with the pid.
+To extend the analysis performable on build graphs, strace \cite{} is used in
+conjunction with our enhanced version of GNU Make.  Strace can capture and record the system
+calls of every process launched during a make build, allowing us to later map directly between
+a shell command launched by make and the system calls it ran.  Strace can write the system call
+information for each pid used to its own file.  This is a good way to map directly from a
+shell command to a set of system calls, except that Linux may reuse pids during the course of
+a build, which causes strace, by default, overwrite these files.  To solve this problem and
+maintain a 1 to 1 mapping between processes launched by make and system call information,
+we modified strace to append a timestamp to each file extension along with the pid.
 
-Once the correct system call data is mapped to the correct build command, the strace data is
-processed to determine which files were inputs and outputs of that command.  An input in this case
-is a file that could have been read but not written and an output is a file that could have been
-written.  Using this information we can modify the build graph and we will discuss this more
+When the system call data is mapped to a command, the data is processed to determine which
+files were inputs and which files were outputs of that command.  An input file could have been
+read by the command but not written, and an output file could have been written by the command.
+Whether or not a file is an input or output depends on which system calls refer to the file.
+Using this information the build graph can potentially be modified and we discuss this more
 in the next section.
@@
 
-The build graph is made up of two types of vertices and two types of directed edges.  A node is
-either a leaf, which represents a file or a single shell command executed by make, or a non-leaf,
-which represents a single target.  An edge is either a dependency edge, which indicates that the
-vertice the edge goes to can run in parallel with other vertices reach by dependency edges from the
-same source vertice, or a recipe edge, which indicates that the vertice the edge points to may not
-run in parallel with other vertices reachable from the source vertice, and it must run in a
-specific order relative to the other vertices reachable from the source.
+A build graph is made up of two types of vertices and two types of directed edges.  A vertex is
+either a leaf, representing a file or a shell command executed by make, or a non-leaf,
+representing a target.  An edge is either a dependency edge, or a recipe edge.
+Dependency edges leaving the same vertex may run in parallel with one another, but must all
+complete before any recipes edges leaving the vertex may be run.  Recipe edges leaving the same
+vertex must run sequentially in a predetermined order after all dependency edges have been run.
 
-To increase the span of the build graph we would like to transform recipe edges into dependency
-edges if possible.  To determine if this is possible we use the collected system call data and
-calculate the file inputs and file outputs of each shell command.  If two adjacent shell commands
-which occur sequentially, share zero or more input files and share no input and output files, then
-those two commands can run in parallel.
+To increase the span of a build graph recipe edges need to transform into dependency
+edges if possible.  To determine if it is possible system call data is used to calculate the
+file inputs and file outputs of each shell command.  If two adjacent recipes share zero or more
+input files and share no input and output files, then those two commands can run in parallel.
 
-Our graph re-builder performs this analysis on the original build graph and produces a new build
-graph which hopefully has a lower span than the original graph.
+For each vertex in the original graph, the inputs and outputs are calculated for each
+dependency and recipe.  And an \emph{input-output} graph is built using this information.  If two
+recipes intersect in more than just their inputs, there is an edge from the recipe that executes
+first to the recipe that executes later.  Every recipe is compared to every other recipe and to
+the set of dependencies to build this \emph{input-output} graph.  The recipes that can be safely
+run in parallel are determined by calculating the lowest level of the graph each vertex is at.
+Recipes may run in parallel with other recipes or dependencies at the same level, but must run
+after all recipes and dependencies at lower levels have completed.  After the new graph is
+constructed the same analysis that was performed on the original graph can be performed on the
+new graph.  Whether or not the new graph has a lower span is dependent upon whether or not
+transformed edges fall within the original span path or not.
 
@@
 \section{Make build analyses}
+\label{sec:realanalyses}
 
-In this section we will look at four problems with Make build systems.  First we will discuss
-what it means for a build to be unnecessarily sequential and how we can diagnose and suggestion
-solutions to that problem.  Second, we will discuss build systems being dominated by a single
-expensive command that prevents the build from running more quickly and how we can diagnose and
-suggest solutions to that problem.  Third, we will discuss when a build has ample available
-parallelism but doesn't achieve its maximum paralleism due to bad operating system scaling.
-Fourth, we will discuss the problem of a slow compiler.
+In this section we discuss four problems with project builds that use make.  In section \ref{sec:expensive},
+we discuss
+builds whose spans are dominated by a single expensive command that prevents the build from
+achieving greater parallelism, look at a project that exhibits this behavior, and explain our
+our tool can diagnose this problem and identify areas for improvement.  In section \ref{sec:sequential}
+, we discuss what
+it means for a build to be unnecessarily sequential, and look at an example of a build that
+exhibits this behavior, and how our tool can diagnose this problem and identify areas for improvement.
+In section \ref{sec:scalability}, we discuss when a build should have ample available parallelism
+but fails to achieve that
+parallelism in practice, look at a project that exhibits this behavior, and discuss potential causes.
+In section \ref{sec:compiler} we discuss another instance of an expensive command dominating a
+build's span; in this case a
+slow compiler, we look at a project that exhibits this behavior and how our tool can diagnose it
+and identify areas for improvement.
 
@@
   \includegraphics[width=0.5\textwidth]{zstd-speedup}
-  \caption{todo}
+  \caption{Speedup of zstd's original build when built with regular make,
+    compared to speedup of zstd's modified build described in section \ref{sec:expensive}
+  when built with regular make.}
   \label{fig:zstd}
@@
 
-All build systems which lack the desired parallelism can be characterized as having too great of
+All project builds which lack the desired parallelism can be characterized as having too great of
 a span in relation to their work.  The span in a build system is the longest (in time) series of
 tasks that must be executed sequentially.  In some cases the span is not a series of similar length
-tasks but is dominated by a single long running task.  In cases like these it is best to point the
-user to the most expensive task in the span because reducing the running time of that task is the
-best way to increase the parallelism of the build.
+tasks, but is dominated by a single long running task.  In cases like these it is best to identify
+the most expensive task in the span, because reducing the running time of that task is the
+most likely way to make progress towards increasing the parallelism of the build.
 
-Reducing the running time of non-span tasks will not decrease the running time of the project as
-a whole.
+To diagnose the problem of a span dominated by a single long running task, our tool can be used
+to learn the span path through the build graph and learn which target and shell commands
+are the most time intensive.  To see what sort of an improvement would occur in the span if the
+running time of the most expensive target in the span was to decrease, the tool will recalculate
+the span of the graph excluding its most expensive member any number of times.  This is useful
+in the case where decreasing the running time of a single task gives the build graph a new span
+path with a similarly long running time, which is dominated by a different long running command.
 
-To diagnose this problem the user can use the tool to discover the span path through the build
-graph and see which target and shell commands are the most time intensive.  The user can then see
-what sort of improvement would occur to the span if they were able to decrease the running
-time of any target in the graph.  This is useful in the case where decreasing the running time
-of a single task gives the build graph a new span path which is dominated by a different also
-long running command that prevents the span of the graph from decreasing.
+Facebook's zstd project \cite{} has a build that exhibits this property.  When we used our tool to analyse
+the span of zstd's build we saw that it is dominated by a single call to cc; see figure \ref{code:cc1}.
 
-An example of a project with such a build is Facebook's zstd project \cite{}.  If we use our tool
-to analyse the span of zstd's build, we see that the original span is dominated by a single call
-to cc.
-
+\begin{figure}[H]
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
+\caption{A long running cc command which was responsible for ??\% of zstd's span.  It compiled and linked 25 .c files.}
+\label{code:cc1}
+\end{figure}
 
-If this call is no longer part of the span, we find that the new span is similarly dominated by
-another single call to cc.
-
+If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
+see figure \ref{code:cc2}.
+\begin{figure}[H]
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
+\caption{A long running cc command which was responsible for ??\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
+This command compiled and linked 30 files.}
+\label{code:cc2}
+\end{figure}
 
-If neither of these calls were members of the build's span, then our tool predicts the span would be
-7.2 seconds rather than 39 seconds or 37 seconds.  To test this we manually replaced the first of
-these two cc calls with 28 new calls to cc; 27 of those are spent compiling each .c file individually
-and the last is to link them all together.  Due to this new organization the 27 .c files can be
-compiled in parallel.  Similarly, the second cc call is replace with 28 new calls to cc; 27 of which
-compile each .c file individually and the last of which links the object files together.
+If neither of these calls are members of the build's span, then our tool predicts the span would be
+approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
+This is about an 80\% reduction in the build's span, and the parallelism improves from 2.95 to ??.
 
-You can see in figure \ref{fig:zstd} that when we ran the modified build with 1 to 32 jobs we
-observed the predicted speedup.  An additional performance improvement that we did not predict was
-that the work of the new build would decrease from approximately 120 seconds to approximately 80 seconds.
+To test this we manually replaced the cc command in figure \ref{code:cc1} with the code in figure
+\ref{code:cc3}.  The original command is replace by a new target for each object file and target
+that can build each object file in parallel and link them together afterwards.
 
-Each of the generated object files is declared as its own target; so when make considers building
-each object file it can check if the file already exists and whether it must be rebuilt.  It turns
-out that the object files being built as part of the first long cc command, had already been
-previously built by make.  By declaring each object file as its own target and building each one
-individually then linking them, make was able to determine it wasn't necessary to build them the
-2nd time; so the effect time of the first cc command when from 40 seconds to 2 seconds; only the
-time necessary for linking the object files together.
+\begin{figure}[H]
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+../lib/common/debug.o: ../lib/common/debug.c                                                         
+    cc \$(FLAGS) ../lib/common/debug.c
+
+\textrm{\emph{26 more targets; one for each object file}}
+
+zstd-build: \$(OBJ\_FILES)
+    cc \$(FLAGS) \$(OBJ\_FILES) -o zstd
+\end{Verbatim}
+\caption{Makefile targets which replaced original code in fig. \ref{code:cc1}}
+\label{code:cc3}
+\end{figure}
+
+Similarly, the code in figure \ref{code:cc2} is replaced by the code in figure \ref{code:cc4}.  Once
+again each object file is explicitly declared as its own target and a target is declared that can
+build these object files in parallel and then link them together afterwards.
+
+\begin{figure}[H]
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+common/debug.o1: common/debug.c
+    cc \$(FLAGS) common/debug.c -o common/debug.o1
+
+\textrm{\emph{29 more targets; one for each object file}}
+
+libzstd-build: \$(OBJ\_FILES)
+    cc \$(FLAGS) \$(OBJ\_FILES) -shared
+       -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+\end{Verbatim}
+\caption{Makefile targets which replaced original code in fig. \ref{code:cc2}}
+\label{code:cc4}
+\end{figure}
+
+Figure \ref{fig:zstd} shows that when the modified build was run with 1 to 32 jobs,
+the predicted speedup was observed.  An additional performance improvement that was not predicted
+is that the work of the new build decreased from approximately 120 seconds to approximately 80 seconds.
+This is because each of the object files is declared as its own target, so make can consider whether
+or not the file already exists and if it is necessary to rebuild it.  In this case, it turns out that
+the object files being built as part of the cc command in fig. \ref{code:cc1} had already been built
+by make; making it unnecessary to build them again.  This cuts the effect time of the code in fig.
+\ref{code:cc1} from about 40 seconds to about 2 seconds; just the time to link the object files.
 
 \subsection{Unnecessarily sequential}
+\label{sec:sequential}
 
@@
 
-\subsection{Bad OS Scaling}
+\subsection{Bad scalability}
+\label{sec:scalability}
 
-\begin{figure}[]
+\begin{figure}[t]
   \includegraphics[width=0.5\textwidth]{llvm-speedup}
-  \caption{todo}
+  \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
+    to the potential speedups our tool predicted using the work and span calculations}
   \label{fig:llvm}
@@
 
-Some projects may exhibit ample potential parallelism, but even when run with more threads than
-available parallelism, predicted parallelism may not be achieved.  LLVM \cite{} is the largest
-project we analysed with a calculated work of approximately 12584 seconds, or 3.5 hours.  The
-predicted span of LLVM is 305 seconds, or 5 minutes.  This gives LLVM a parallelism of about 41.
-These calculations where made from a build of LLVM that was done not using strace.
+Some projects may exhibit ample potential parallelism but may not achieve that potential
+parallelism in practice.  LLVM \cite{} is the largest
+project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
+With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
+These calculations where made from a build of LLVM that was run without strace.
 
-A parallelism of 41 would lead one to believe that on a machine with 32 cores LLVM should be built
-in parallel in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
+A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
+LLVM should build in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
 built with ``-j 32'' the observed speedup was slightly over 18; see chart \ref{}.  Using work and
@@
 \subsection{Slow compiler}
+\label{sec:compiler}
 
-A similar issue to the one described in section \ref{sec:expensive} is when a build's
-span is dominated by the compilation of a single file.  This problem however cannot be
-solved in the makefile like the zstd issue was.  Solutions to this problem include
+An issue similar to the one described in section \ref{sec:expensive} is a build's
+span being dominated by the compilation of a single file.  This problem cannot be
+solved in the makefile like the issue in section \ref{sec:expensive} was.
+Possible solutions to this problem include
 breaking the file up into smaller files which can be compiled in parallel and then
-linked; or changing the compiler so that it is capable of compiling the file faster.
-A faster compiler, would have the additional benefit of not only decreasing the span
+linked, or changing the compiler so it is capable of compiling the file faster.
+A faster compiler would likely have the additional benefit of not only decreasing the span
 of the project but also likely decreasing the work of the project.
@@
 Google's protocol buffers project \cite{} exhibited this behavior.  It had a relatively
-long predicted span of ?? seconds, which was dominated by a call to g++ which was
-compiling a single file.  When we apply the same analysis as we applied to zstd in
-section \ref{sec:expensive}, we see that if \ref{} wasn't a part of the span, the new
-span would be ?? and would be dominated by another g++ compilation of a single file.
-And we continue this again to find that the 3 top span paths are dominated by g++
-compilations of individual files.  If none of these calls were part of the span, the
-new span of the project would be ??.  Now if the developers of protocol buffers were
-interested in increasing the parallelism of their project they would know where they
-should focus their efforts
+long predicted span of ?? seconds, which was dominated by a call to g++ which compiles
+a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
+protocol buffers' build, we learn that if the code in fig. \ref{code:g++1} wasn't part of the span,
+the new span would be ?? and would be dominated by the code in fig. \ref{code:g++2}.  And when that
+code isn't part of the span, we learn that the new span is again dominated by another
+g++ call in the code in fig. \ref{g++3}.  If none of these g++ compilations are in the span,
+the new span of the project is roughly half the original span.
 
-Here is a simplification of the real command run by make; it took about 40 seconds to run.
+If the developers of protocol buffers were interested in increasing the parallelism of
+their project they would know they should begin their efforts by speeding up the
+compilation of these 3 files.
+
+\begin{figure}[H]
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
+\caption{A simplification of the real command run by make; compiles a single file and took about
+40 seconds to run.}
+\label{code:g++1}
+\end{figure}
 
-Here is a simplification of the real command run by make; it took about 27 seconds to run.
+\begin{figure}[H]
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
+\caption{A simplification of the real command run by make; compiles a single file and took about
+27 seconds to run.}
+\label{code:g++2}
+\end{figure}
 
-Here is a simplification of the real command run by make; it took about 22.5 seconds to run.
+\begin{figure}[H]
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
+\caption{A simplification of the real command run by make; compiles a single file and took about
+22.5 seconds to run.}
+\label{code:g++3}
+\end{figure}
 
@@
 
-One of the limitations of this work is that currenlty to collect the necessary data to rebuild
+One of the limitations of this work is that currenlty to collect the necessary data to reconstruct
 the build graph, the project must be built using a single thread.  This is a burden for projects
-that are very long running such as LLVM.  Additionally, building with strace causes an additional
+that are very long running, such as LLVM.  Additionally, building with strace causes an additional
 slowdown and inflates the running time of each shell command, which makes the raw numbers
@@
 it would be good to collect the necessary information while allowing make to run in parallel with
-any number of processes.
+any number of processes.  It would also be good to know how much strace perturbs the recorded
+time of any given shell command so both the tool with and without strace don't need to be run.
 

### 2019-01-17 paper.tex
Citations.
They finally told us how to format.

@@
 
-\documentclass[sigplan,10pt,review,authorversion]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-
+\documentclass[sigconf,10pt,review,authorversion]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\usepackage{geometry}
+\geometry{reset, twoside=true, head=13pt,
+     paperwidth=8.5in, paperheight=11in,
+     includeheadfoot, columnsep=2pc,
+     top=57pt, bottom=73pt, inner=72pt, outer=72pt,
+     marginparwidth=2pc,heightrounded
+     }
 \citestyle{acmnumeric}
@@
 Some projects may exhibit ample potential parallelism but may not achieve that potential
-parallelism in practice.  LLVM \cite{} is the largest
+parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
 project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
@@
 
+\bibliographystyle{ACM-Reference-Format} 
+\bibliography{citations}
+\balance
+
 \end{document}

### 2019-01-17 paper.tex
committing to pull

--- 2019-01-17 paper.tex
@@
 to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
-discuss the modified version of GNU Make \cite{} we use to record enough information about a
+discuss the modified version of GNU Make \cite{gnumake4.2} we use to record enough information about a
 project's build to reconstruct the build graph during a post-mortem analysis.
@@
 
-To collect the desired level of information from a run of GNU Make \cite{}, we run it with the
+To collect the desired level of information from a run of GNU Make \cite{gnumake4.2}, we run it with the
 debug level of 'verbose', use our own wrappers, which print additional information, and use
@@
 
-To extend the analysis performable on build graphs, strace \cite{} is used in
+To extend the analysis performable on build graphs, strace \cite{strace4.25} is used in
 conjunction with our enhanced version of GNU Make.  Strace can capture and record the system
@@
 
-Facebook's zstd project \cite{} has a build that exhibits this property.  When we used our tool to analyse
-the span of zstd's build we saw that it is dominated by a single call to cc; see figure \ref{code:cc1}.
+Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyse
+the span of zstd's build we saw that it is dominated by a single call to cc; see fig. \ref{code:cc1}.
 
@@
 \end{Verbatim}
-\caption{A long running cc command which was responsible for ??\% of zstd's span.  It compiled and linked 25 .c files.}
+\caption{A long running cc command which was responsible for about 97\% of zstd's span.  It compiled and linked 27 .c files.}
 \label{code:cc1}
@@
 If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
-see figure \ref{code:cc2}.
+see fig. \ref{code:cc2}.
 \begin{figure}[H]
@@
   \textrm{\emph{28 more .c files}}
-  -shared -fPIC -fvisibility=hidden
   -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
 \end{Verbatim}
-\caption{A long running cc command which was responsible for ??\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
+\caption{A long running cc command which was responsible for aproximately 100\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
 This command compiled and linked 30 files.}
@@
 approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
-This is about an 80\% reduction in the build's span, and the parallelism improves from 2.95 to ??.
+This is about an 80\% reduction in the build's span, and the parallelism improves from 2.98 to 16.
 
-To test this we manually replaced the cc command in figure \ref{code:cc1} with the code in figure
+To test this we manually replaced the cc command in fig. \ref{code:cc1} with the code in fig.
 \ref{code:cc3}.  The original command is replace by a new target for each object file and target
@@
 
-Similarly, the code in figure \ref{code:cc2} is replaced by the code in figure \ref{code:cc4}.  Once
+Similarly, the code in fig. \ref{code:cc2} is replaced by the code in fig. \ref{code:cc4}.  Once
 again each object file is explicitly declared as its own target and a target is declared that can
@@
 
-Makefiles are a collection of rules which describe how to build some targets.  Each rule consists
+Makefiles are a collection of rules which describe how to build targets \cite{gnumakemanual}.
+Each rule consists
 of a a series of prerequisites and a series of recipes which are run to build the target.  The
-prerequisites can be built in parallel but the recipes are run in the order they are listed.  So,
-the author of a makefile can affect the potential parallelism of the build by how they write their
-makefile.
+prerequisites can be built in parallel, but the recipes must be run in the order they are listed.  So,
+the author of a makefile can directly impact the potential parallelism of the build by
+how they organize their makefile.
 
-A makefile which is too sequential will have a longer than necessary span.  In the previous
-section we talked about projects having a long span because it was dominated by a single
-long running shell command.  Another possiblility is the span being made up of many shell
-commands; none of which dominate.  In the previous section we talked about zstd, and if we
-rewrite the makefiles so the long running commands are made up of many sequential compilation
-and then linking commands we end up with a build that has the same work and span as the original.
+In the previous section we discussed zstd's build, which turned out to be unnecessarily sequential
+because two single shell calls which compiled and linked many files at once could be broken up
+into many parallel compilation jobs which could then be linked after.  These two commands could
+have been written as follows in fig. \ref{code:cc5} and fig. \ref{code:cc6} respectively, instead of as they were
+originally in fig. \ref{code:cc1} and fig. \ref{code:cc2} and would have resulted in the same
+build performance.
+
+\begin{figure}[H]
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+cc \$(FLAGS) ../lib/common/debug.c
+   -o ../lib/common/debug.o
+cc \$(FLAGS) ../lib/common/entropy\_common.c
+   -o ../lib/common/entropy\_common.o
+\textrm{\emph{25 more recipes; one for each .c file that needs to be compiled}}
+cc \$(FLAGS) \$(OBJ\_FILES) -o zstd
+\end{Verbatim}
+  \caption{An alternative to the code in fig. \ref{code:cc1}. Each file is compiled individually
+    as its own recipe, followed by a recipe to link the object files together.  This has the same
+    performance as the original command.}
+\label{code:cc5}
+\end{figure}
+
+\begin{figure}[H]
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+cc \$(FLAGS) common/debug.c -o common/debug.o1
+cc \$(FLAGS) common/entropy\_common.c
+   -o common/\_common.o1
+\textrm{\emph{28 more recipes; one for each .c file that needs to be compiled}}
+cc \$(FLAGS) \$(OBJ\_FILES) -shared
+   -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+\end{Verbatim}
+  \caption{An alternative to the code in fig. \ref{code:cc2}. Each file is compiled individually
+    as its own recipe, followed by a recipe to link the object files together.  This has the same
+    performance as the original command.}
+\label{code:cc6}
+\end{figure}
 
@@
 \begin{figure}[t]
-  \includegraphics[width=0.5\textwidth]{llvm-speedup}
+  \includegraphics[width=0.5\textwidth]{llvm-speedup-w-64}
   \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
@@
 Some projects may exhibit ample potential parallelism but may not achieve that potential
-parallelism in practice.  LLVM \cite{} is the largest
+parallelism in practice.  LLVM \cite{llvm6.0.0} is the largest
 project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
@@
 
-Google's protocol buffers project \cite{} exhibited this behavior.  It had a relatively
-long predicted span of ?? seconds, which was dominated by a call to g++ which compiles
+Google's protocol buffers project \cite{protobufs3.6.1} exhibited this behavior.  It's work was calculated
+to be about 16.7 minutes and it had a relatively
+long predicted span of 43.7 seconds, which was dominated by a call to g++ compiling
 a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
 protocol buffers' build, we learn that if the code in fig. \ref{code:g++1} wasn't part of the span,
-the new span would be ?? and would be dominated by the code in fig. \ref{code:g++2}.  And when that
+the new span would be about 31s and would be dominated by the code in fig. \ref{code:g++2}.
+And when that
 code isn't part of the span, we learn that the new span is again dominated by another
 g++ call in the code in fig. \ref{g++3}.  If none of these g++ compilations are in the span,
-the new span of the project is roughly half the original span.
+the new span of the project is about 20s, roughtly half of the original span.
 
@@
 \caption{A simplification of the real command run by make; compiles a single file and took about
-40 seconds to run.}
+40 seconds to run and was 90\% of the span.}
 \label{code:g++1}
@@
 \caption{A simplification of the real command run by make; compiles a single file and took about
-27 seconds to run.}
+27 seconds to run and was about 87\% of the span.}
 \label{code:g++2}
@@
 \caption{A simplification of the real command run by make; compiles a single file and took about
-22.5 seconds to run.}
+22.5 seconds to run and was about 90\% of the span.}
 \label{code:g++3}
@@
 
-One of the limitations of this work is that currenlty to collect the necessary data to reconstruct
+One of the limitations of this work is that currently to collect the necessary data to reconstruct
 the build graph, the project must be built using a single thread.  This is a burden for projects
@@
 
+\bibliography{citations}
+\bibliographystyle{plain}
+
 \end{document}

### 2019-01-17 paper.tex

@@
 Some projects may exhibit ample potential parallelism but may not achieve that potential
-parallelism in practice.  LLVM \cite{llvm6.0.0} is the largest
+parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
 project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
@@
 
+\bibliographystyle{ACM-Reference-Format} 
 \bibliography{citations}
-\bibliographystyle{plain}
+\balance
 

### 2019-01-17 paper.tex
Finish intro.

--- 2019-01-17 paper.tex
@@
 degrees of parallelism.
+
+To address this problem, numerous new build systems have been created
+(Bazel, Buck, Pants, Aegis, Shake, CMake, Ninja, ...), many purporting
+to solve or at least alleviate these problems. However, systems
+software ranging from the Linux kernel to LLVM to zlib is still built
+with \verb|make| today, suggesting that as in many areas, building new
+tools is not sufficient---we must analyze, understand, and improve use
+of existing tools as well.
+
+To accomplish this for build systems, we have developed a tool for
+analyzing \verb|make|-based builds by instrumenting them while they
+run, relying on \verb|make|'s own internal information and the use of
+\verb|strace| to analyze the behavior of the actual commands used to
+run the build. Our tool computes a dependence graph based on the
+actual behavior of the build, then prunes it with data from
+\verb|strace|. The resulting graph can accurately predict the
+potential parallelism of the build from the instrumentation, show what
+build steps cause the biggest slowdown, and answer counterfactual
+questions about the results of improving the build system.
+
+After describing our analysis tool, we present four case studies in
+reasons software might be slow to build, backed by data and analysis
+from our tool. All of the case studies are real, widely-used
+open-source projects. In some cases, we identify low-hanging fruit
+that can increase the parallelism of a build by 4x; in others we point
+to places where code organization changes could make a significant
+difference. These case studies demonstrate the benefits of a
+build-system analysis tool and point to valuable further investigation.
 

### 2019-01-17 paper.tex
to merge

--- 2019-01-17 paper.tex
@@
 a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
-protocol buffers' build, we learn that if the code in fig. \ref{code:g++1} wasn't part of the span,
-the new span would be about 31s and would be dominated by the code in fig. \ref{code:g++2}.
+protocol buffers' build, we learn that if the code, a simplification of the real code,
+in fig. \ref{code:g++1} wasn't part of the span,
+the new span would be about 31s and would be dominated by the code in fig. \ref{code:g++2}, also
+a simplification.
 And when that
 code isn't part of the span, we learn that the new span is again dominated by another
-g++ call in the code in fig. \ref{g++3}.  If none of these g++ compilations are in the span,
+g++ call in the code in fig. \ref{g++3}, also a simplification.
+If none of these g++ compilations are in the span,
 the new span of the project is about 20s, roughtly half of the original span.
@@
 \end{Verbatim}
-\caption{A simplification of the real command run by make; compiles a single file and took about
-40 seconds to run and was 90\% of the span.}
+\caption{Compiles a single file and took about 40 seconds to run and was 90\% of the span.}
 \label{code:g++1}
@@
 \end{Verbatim}
-\caption{A simplification of the real command run by make; compiles a single file and took about
-27 seconds to run and was about 87\% of the span.}
+\caption{Compiles a single file and took about 27 seconds to run and was about 87\% of the span.}
 \label{code:g++2}
@@
 \end{Verbatim}
-\caption{A simplification of the real command run by make; compiles a single file and took about
-22.5 seconds to run and was about 90\% of the span.}
+\caption{Compiles a single file and took about 22.5 seconds to run and was about 90\% of the span.}
 \label{code:g++3}

### 2019-01-17 paper.tex
rewrite conclusion

--- 2019-01-17 paper.tex
@@
 happy coincidence of software complexity and hardware trends.
-
+%
 Unfortunately, the story is not quite this rosy, as almost anyone who
@@
 compiling large software projects use highly parallel systems to do so
-[cite Regehr tweet].
+\cite{regehr-tweet}.
 
@@
 highly-parallel systems, real software rarely makes use of the full
-capacity of the machine it is built on, despite the potential for high
+capacity of the machine despite the potential for high
 degrees of parallelism.
@@
 
-\section{Future work}
+\section{Conclusion}
 
-One of the limitations of this work is that currently to collect the necessary data to reconstruct
-the build graph, the project must be built using a single thread.  This is a burden for projects
-that are very long running, such as LLVM.  Additionally, building with strace causes an additional
-slowdown and inflates the running time of each shell command, which makes the raw numbers
-reported by a build graph created with strace on, inaccurate.  In the future for performance sake
-it would be good to collect the necessary information while allowing make to run in parallel with
-any number of processes.  It would also be good to know how much strace perturbs the recorded
-time of any given shell command so both the tool with and without strace don't need to be run.
+We present an automated analysis tool for build systems based on
+\verb|make|, which can automatically determine the available
+parallelism in a build process and provide useful information to
+improve it. Our tool makes use of \verb|strace| instrumentation to
+determine the ground truth about the build process, and can answer
+counterfactual questions about how much a build could improve.
+%
+We show four case studies on three real software projects
+demonstrating the power of our analysis and point to sources of
+slowdown and lack of parallelism, and use our tool to make
+improvements that substantially increase parallelism.
 
-In addition to this, we hope to analyse more project builds and introduce parallelism at other
-level of project builds, such as at the level of the compiler.
+The case studies we've shown already demonstrate the value of our
+tool, but extensions could make it more useful to developers and build
+system authors---for example, by automatically generating revised
+\verb|Makefile|s or by using \verb|strace| to instrument arbitrary
+build systems, not just \verb|make|. Finally, we hope to integrate
+this with an analysis of parallelism in commands executed by the build
+system, such as the \verb|gold| linker.
 

### 2019-01-17 paper.tex
shrunk some figures

--- 2019-01-17 paper.tex
@@
   ../lib/common/debug.c
-  ../lib/common/entropy\_common.c
-  \textrm{\emph{25 more .c files}}
+  \textrm{\emph{26 more .c files}}
   zstdcli.o fileio.o bench.o
@@
   common/debug.c
-  common/entropy\_common.c
-  \textrm{\emph{28 more .c files}}
+  \textrm{\emph{29 more .c files}}
   -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
@@
     cc \$(FLAGS) ../lib/common/debug.c
-
 \textrm{\emph{26 more targets; one for each object file}}    
-
 zstd-build: \$(OBJ\_FILES)
@@
 common/debug.o1: common/debug.c
-    cc \$(FLAGS) common/debug.c -o common/debug.o1
-
+    cc \$(FLAGS) common/debug.c
+       -o common/debug.o1
 \textrm{\emph{29 more targets; one for each object file}}
-
 libzstd-build: \$(OBJ\_FILES)
     cc \$(FLAGS) \$(OBJ\_FILES) -shared
-       -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+       -o libzstd.so.1.3.7
 \end{Verbatim}
@@
    -o ../lib/common/debug.o
-cc \$(FLAGS) ../lib/common/entropy\_common.c
-   -o ../lib/common/entropy\_common.o
-\textrm{\emph{25 more recipes; one for each .c file that needs to be compiled}}
+\textrm{\emph{26 more recipes; one for each .c file that needs to be compiled}}
 cc \$(FLAGS) \$(OBJ\_FILES) -o zstd
@@
 cc \$(FLAGS) common/debug.c -o common/debug.o1
-cc \$(FLAGS) common/entropy\_common.c
-   -o common/\_common.o1
-\textrm{\emph{28 more recipes; one for each .c file that needs to be compiled}}
+\textrm{\emph{29 more recipes; one for each .c file that needs to be compiled}}
 cc \$(FLAGS) \$(OBJ\_FILES) -shared
-   -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+   -o libzstd.so.1.3.7
 \end{Verbatim}
@@
 long predicted span of 43.7 seconds, which was dominated by a call to g++ compiling
-a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
-protocol buffers' build, we learn that if the code, a simplification of the real code,
-in fig. \ref{code:g++1} wasn't part of the span,
-the new span would be about 31s and would be dominated by the code in fig. \ref{code:g++2}, also
-a simplification.
-And when that
-code isn't part of the span, we learn that the new span is again dominated by another
-g++ call in the code in fig. \ref{g++3}, also a simplification.
-If none of these g++ compilations are in the span,
-the new span of the project is about 20s, roughtly half of the original span.
-
-If the developers of protocol buffers were interested in increasing the parallelism of
-their project they would know they should begin their efforts by speeding up the
-compilation of these 3 files.
+a single file.  When the analysis applied in section \ref{sec:expensive} is applied
+to protocol buffers' build, we learn that the top three spans are dominated by the
+single file compilations of the files in fig. \ref{code:g++1}.  If the g++ compilations
+of these three files were not in the build's span the new span would be about 20s,
+roughly half of the original span.  Speeding up the compilation of these three files
+would be a good place to begin improving the parallelism of this build.
 
@@
 \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  g++ \$(FLAGS)
-  -o google/protobuf/descriptor.lo
   google/protobuf/descriptor.cc
-\end{Verbatim}
-\caption{Compiles a single file and took about 40 seconds to run and was 90\% of the span.}
-\label{code:g++1}
-\end{figure}
-
-\begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  g++ \$(FLAGS)
-  -o google/protobuf/descriptor.pb.lo
   google/protobuf/descriptor.pb.cc
-\end{Verbatim}
-\caption{Compiles a single file and took about 27 seconds to run and was about 87\% of the span.}
-\label{code:g++2}
-\end{figure}
-
-\begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  g++ \$(FLAGS)
-  -o google/protobuf/compiler/cpp/cpp\_message.lo
   google/protobuf/compiler/cpp/cpp\_message.cc
 \end{Verbatim}
-\caption{Compiles a single file and took about 22.5 seconds to run and was about 90\% of the span.}
-\label{code:g++3}
+\caption{The compilation of these 3 files dominate the longest 3 spans.}
+\label{code:g++1}
 \end{figure}

### 2019-01-17 paper.tex
cites

@@
 to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
-discuss the modified version of GNU Make \cite{gnumake4.2} we use to record enough information about a
+discuss the modified version of GNU Make \cite{gnumakemanual} we use to record enough information about a
 project's build to reconstruct the build graph during a post-mortem analysis.
@@
 
-To collect the desired level of information from a run of GNU Make \cite{gnumake4.2}, we run it with the
+To collect the desired level of information from a run of GNU Make \cite{gnumakemanual}, we run it with the
 debug level of 'verbose', use our own wrappers, which print additional information, and use

### 2019-01-17 paper.tex
fix missing ref
missing cite
work on unnecessarily sequential section

--- 2019-01-17 paper.tex
@@
 \end{figure}
-
 If neither of these calls are members of the build's span, then our tool predicts the span would be
@@
 
-Makefiles are a collection of rules which describe how to build targets \cite{gnumakemanual}.
-Each rule consists
+Makefiles are a collection of rules which describe how to build targets \cite{gnumakemanual},
+and each rule consists
 of a a series of prerequisites and a series of recipes which are run to build the target.  The
@@
 the author of a makefile can directly impact the potential parallelism of the build by
-how they organize their makefile.
+how they organize their makefile.  It is possible to unnecessarily list things as recipes that
+could have run in parallel.  Here we will look at an example of this and how our tool can
+discover that these recipes don't actually need to run in parallel.
 
@@
 originally in fig. \ref{code:cc1} and fig. \ref{code:cc2} and would have resulted in the same
-build performance.
+build performance, but would make it possible for our tool to detect that they are unnecessarily
+sequential.
 
-\begin{figure}[H]
+The following code is an alternative to the code in fig. \ref{code:cc1}. Each file is compiled individually
+as its own recipe, followed by a recipe to link the object files together.  This has the same
+performance as the original command.
   \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
-  \caption{An alternative to the code in fig. \ref{code:cc1}. Each file is compiled individually
+
+The following code is an alternative to the code in fig. \ref{code:cc2}. Each file is compiled individually
   as its own recipe, followed by a recipe to link the object files together.  This has the same
-    performance as the original command.}
-\label{code:cc5}
-\end{figure}
-
-\begin{figure}[H]
+  performance as the original command.
   \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
@@
 \end{Verbatim}
-  \caption{An alternative to the code in fig. \ref{code:cc2}. Each file is compiled individually
-    as its own recipe, followed by a recipe to link the object files together.  This has the same
-    performance as the original command.}
-\label{code:cc6}
-\end{figure}
 
-To diagnose this problem using our tool, the user can once again calculate the span of the build
-graph and look at what targets and shell calls constitute the span.  In this case the span is
-made up of many short running compilation jobs.  To check if these compilation jobs may run in
-parallel, our tool will take the original graph, the collected strace data, and build a new graph
-as described in section \ref{sec:revisegraphs}.  The span of the new graph can then be compared
-to the span of the old graph; if the span of the new graph is less than the span of the old graph,
-then the original graph was unnecessarily sequential.  The allow this build to run faster in the
-future, the developer can rewrite the makefile to reflect the organization of the build in
-the new graph.
-
-In the case of zstd, when built with strace; the calculated work for the original graph is 133
-seconds and the span is 45 seconds; which means a parallelism of 2.95.  When this graph is
-rebuilt using strace data and the span is recalculated, it is 8 seconds, which matches with the
-predictions and observations from section \ref{sec:expensive}.
+  These recipes compiling individual files can run in parallel, as shown in the previous section,
+  and our tool can detect this by using the build graph and the system call information gathered
+  by strace and applying the methods described in section \ref{revisegraphs}.  When the span
+  of the new graph is calculated it reflects the fact that these compilations can occur in
+  parallel and is 8 seconds rather than the original 45 seconds; these times are longer than
+  those mentioned in the previous section because strace inflates the running time of the
+  commands it traces.
 

### 2019-01-17 paper.tex

--- 2019-01-17 paper.tex
@@
 LLVM should build in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
-built with ``-j 32'' the observed speedup was slightly over 18; see chart \ref{}.  Using work and
+built with ``-j 32'' the observed speedup was slightly over 18; see chart~\ref{fig:llvm}.  Using work and
 span our tool can predict the potential speedup a build will experience using Brent's theorem
-\cite{}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
+\cite{brents}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
 P processors, \begin{math} T_P \leq T \end{math}

### 2019-01-17 paper.tex
fix some reference errors

@@
 into many parallel compilation jobs which could then be linked after.  These two commands could
-have been written as follows in fig. \ref{code:cc5} and fig. \ref{code:cc6} respectively, instead of as they were
-originally in fig. \ref{code:cc1} and fig. \ref{code:cc2} and would have resulted in the same
+have been written as follows, respectively, instead of as they were
+originally in fig. \ref{code:cc1} and fig. \ref{code:cc2}, and would have resulted in the same
 build performance, but would make it possible for our tool to detect that they are unnecessarily
@@
   and our tool can detect this by using the build graph and the system call information gathered
-  by strace and applying the methods described in section \ref{revisegraphs}.  When the span
+  by strace and applying the methods described in section \ref{sec:revisegraphs}.  When the span
   of the new graph is calculated it reflects the fact that these compilations can occur in

### 2019-01-17 paper.tex
format checker doesn't like column numbers.
work on scalability section

@@
 
-Some projects may exhibit ample potential parallelism but may not achieve that potential
+Some projects may exhibit ample potential parallelism but not achieve that potential
 parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
-project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
-With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
-These calculations where made from a build of LLVM that was run without strace.
+project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours, and
+a predicted span of 305 seconds, or 5 minutes, measured from a build of LLVM that was run without
+strace. With a work of 3.5 hours and a span of 5 minutes, LLVM has a parallelism of about 41.
 
 A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
-LLVM should build in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
-built with ``-j 32'' the observed speedup was slightly over 18; see chart~\ref{fig:llvm}.  Using work and
-span our tool can predict the potential speedup a build will experience using Brent's theorem
-\cite{brents}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
-P processors, \begin{math} T_P \leq T \end{math}
+LLVM would build in about 5 minutes.  Our observations however, did not match this expectation.
+In chart \ref{fig:llvm} we see two different observed speedups, indicted by the squares and
+triangles.  The squares indicate the observed speedup of LLVM when run on a machine with 32 cores,
+with \emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the observed speedup when run
+on a different machine, with 32 real cores and 32 hyperthreads, with \emph{make -j n} from where
+n is most of the numbers between 1 and 32.  The dashed line indicates the upper predicted speedup,
+which is the speedup the project can't exceed.  It is apparent from the chart that LLVM fails to
+achieve this maximum predicted speedup.  The solid line represents the predicted lower speedup,
+which is the speedup the project should achieve at minimum.  And it appears in the chart that
+LLVM narrowly achieves this speedup, when built with \emph{make -j n} for most n.
+
+The lower and upper speedup bounds are calculated using Brent's theorem \cite{brents}, which
+specifies upper and lower bounds on the time, $T_P$ it will take a parallel algorithm to run on P
+processors.
+\begin{math}
+  \frac{work}{P} \geq T_P \leq \frac{work}{P} + span 
+\end{math}
 

### 2019-01-17 paper.tex
change math to equation
figures.

--- 2019-01-17 paper.tex
@@
 
-\documentclass[sigconf,10pt,review,authorversion]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigconf,10pt,authorversion]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 \usepackage{geometry}
@@
 \begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  cc \$(FLAGS)
-  ../lib/common/debug.c
-  \textrm{\emph{26 more .c files}}
-  zstdcli.o fileio.o bench.o
-  datagen.o dibio.o -o zstd
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
+  cc \$(FLAGS) common/debug.c \textrm{\emph{26 more .c files}}
+    zstdcli.o fileio.o bench.o datagen.o dibio.o
+    -o zstd
 \end{Verbatim}
@@
 \begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-  cc \$(FLAGS)
-  common/debug.c
-  \textrm{\emph{29 more .c files}}
-  -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
+  cc \$(FLAGS) common/debug.c \textrm{\emph{29 more .c files}}
+    -o libzstd.so.1.3.7
 \end{Verbatim}
-\caption{A long running cc command which was responsible for aproximately 100\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
+\caption{A long running cc command which was responsible for aproximately 100\% of zstd's span when the command in fig. \ref{code:cc1} is removed.
 This command compiled and linked 30 files.}
@@
 \begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-../lib/common/debug.o: ../lib/common/debug.c                                                       
-    cc \$(FLAGS) ../lib/common/debug.c
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
+common/debug.o: common/debug.c                                                       
+    cc \$(FLAGS) common/debug.c
 \textrm{\emph{26 more targets; one for each object file}}    
@@
 \begin{figure}[H]
-  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
 common/debug.o1: common/debug.c
@@
 performance as the original command.
-  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
-cc \$(FLAGS) ../lib/common/debug.c
-   -o ../lib/common/debug.o
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
+cc \$(FLAGS) common/debug.c
+   -o common/debug.o
 \textrm{\emph{26 more recipes; one for each .c file that needs to be compiled}}
@@
   performance as the original command.
-  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
 cc \$(FLAGS) common/debug.c -o common/debug.o1
@@
 \begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
+\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
   google/protobuf/descriptor.cc

### 2019-01-17 paper.tex

@@
 
-Some projects may exhibit ample potential parallelism but may not achieve that potential
+Some projects may exhibit ample potential parallelism but not achieve that potential
 parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
-project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
-With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
-These calculations where made from a build of LLVM that was run without strace.
+project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours, and
+a predicted span of 305 seconds, or 5 minutes, measured from a build of LLVM that was run without
+strace. With a work of 3.5 hours and a span of 5 minutes, LLVM has a parallelism of about 41.
 
 A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
-LLVM should build in about 5 minutes.  Our observations did not match this expectation.  When LLVM was
-built with ``-j 32'' the observed speedup was slightly over 18; see chart~\ref{fig:llvm}.  Using work and
-span our tool can predict the potential speedup a build will experience using Brent's theorem
-\cite{brents}.  The lower speedup bound is predicted using Brent's theorem, which specifies that for
-P processors, \begin{math} T_P \leq T \end{math}
+LLVM would build in about 5 minutes.  Our observations however, did not match this expectation.
+In chart \ref{fig:llvm} we see two different observed speedups, indicted by the squares and
+triangles.  The squares indicate the observed speedup of LLVM when run on a machine with 32 cores,
+with \emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the observed speedup when run
+on a different machine, with 32 real cores and 32 hyperthreads, with \emph{make -j n} from where
+n is most of the numbers between 1 and 32.  The dashed line indicates the upper predicted speedup,
+which is the speedup the project can't exceed.  It is apparent from the chart that LLVM fails to
+achieve this maximum predicted speedup.  The solid line represents the predicted lower speedup,
+which is the speedup the project should achieve at minimum.  And it appears in the chart that
+LLVM narrowly achieves this speedup, when built with \emph{make -j n} for most n.
+
+The lower and upper speedup bounds are calculated using Brent's theorem \cite{brents}, which
+specifies upper and lower bounds on the time, $T_P$ it will take a parallel algorithm to run on P
+processors.
+
+\begin{equation*}
+  \frac{work}{P} \geq T_P \leq \frac{work}{P} + span 
+\end{equation*}
 

### 2019-01-17 paper.tex
fix overfull

@@
 
-This is perhaps best demonstrated by the classic \verb|make -j|
+This is perhaps best demonstrated by the  \verb|make -j|
 command, which builds on \verb|make|~\cite{make} to execute
@@
 to solve or at least alleviate these problems. However, systems
-software ranging from the Linux kernel to LLVM to zlib is still built
+ ranging from the Linux kernel to LLVM to zlib are still built
 with \verb|make| today, suggesting that as in many areas, building new

### 2019-01-17 paper.tex
moved a citation to end of line

@@
 
-Makefiles are a collection of rules which describe how to build targets \cite{gnumakemanual},
+Makefiles are a collection of rules which describe how to build targets,
 and each rule consists
-of a a series of prerequisites and a series of recipes which are run to build the target.  The
+of a a series of prerequisites and a series of recipes which are run to build the target \cite{gnumakemanual}.  The
 prerequisites can be built in parallel, but the recipes must be run in the order they are listed.  So,

### 2019-01-17 paper.tex
llvm section

--- 2019-01-17 paper.tex
@@
 
-\subsection{Bad scalability}
+\subsection{Poor scaling}
 \label{sec:scalability}
@@
   \includegraphics[width=0.5\textwidth]{llvm-speedup-w-64}
-  \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
-    to the potential speedups our tool predicted using the work and span calculations}
+  \caption{A comparison of the observed speedup when LLVM was built
+    with \emph{make -j <n> for <n> from 1 to 64}, to the potential
+    speedups our tool predicted using the work and span calculations}
   \label{fig:llvm}
@@
 
-Some projects may exhibit ample potential parallelism but not achieve that potential
-parallelism in practice.  LLVM~\citep{LLVM:CGO04} is the largest
-project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours, and
-a predicted span of 305 seconds, or 5 minutes, measured from a build of LLVM that was run without
-strace. With a work of 3.5 hours and a span of 5 minutes, LLVM has a parallelism of about 41.
+Some projects may exhibit ample potential parallelism but not achieve
+that potential parallelism in practice.  LLVM~\citep{LLVM:CGO04} is
+the largest project we analysed, with a calculated work of
+approximately 12584 seconds, or 3.5 hours, and a predicted span of 305
+seconds, or 5 minutes (\verb|strace| instrumentation was disabled for
+this measurement). With a work of 3.5 hours and a span of 5 minutes,
+LLVM has a parallelism of about 41.
 
-A parallelism of 41 lead us to believe that when built in parallel on a machine with 32 cores,
-LLVM would build in about 5 minutes.  Our observations however, did not match this expectation.
-In chart \ref{fig:llvm} we see two different observed speedups, indicted by the squares and
-triangles.  The squares indicate the observed speedup of LLVM when run on a machine with 32 cores,
-with \emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the observed speedup when run
-on a different machine, with 32 real cores and 32 hyperthreads, with \emph{make -j n} from where
-n is most of the numbers between 1 and 32.  The dashed line indicates the upper predicted speedup,
-which is the speedup the project can't exceed.  It is apparent from the chart that LLVM fails to
-achieve this maximum predicted speedup.  The solid line represents the predicted lower speedup,
-which is the speedup the project should achieve at minimum.  And it appears in the chart that
-LLVM narrowly achieves this speedup, when built with \emph{make -j n} for most n.
+A parallelism of 41 lead us to believe that when built in parallel on
+a machine with 32 cores, LLVM would build in not that much more than 5
+minutes.  Our observations however, did not match this expectation.
+In chart \ref{fig:llvm} we see two different observed speedups,
+indicted by the squares and triangles.  The squares indicate the
+observed speedup of LLVM when run on a machine with 32 cores, with
+\emph{make -j 1} to \emph{make -j 32}.  The triangles indicate the
+observed speedup when run on a different machine, with 32 cores each
+with 2 hyperthreads, with \emph{make -j n} from where $n$ between 1
+and 64.  The dashed line and solid lines indicates the upper and lower
+bounds we would expect for parallel speedup, calculated using Brent's
+bound~\cite{brents}.  It
+is apparent from the chart that LLVM is consistently along the lower
+bound.
 
-The lower and upper speedup bounds are calculated using Brent's theorem \cite{brents}, which
-specifies upper and lower bounds on the time, $T_P$ it will take a parallel algorithm to run on P
-processors.
-
-\begin{equation*}
-  \frac{work}{P} \geq T_P \leq \frac{work}{P} + span 
-\end{equation*}
+This demonstrates that our automatic prediction is accurate out to a
+large number of parallel jobs, but also that the LLVM build fails to
+achieve as much parallel scaling as we would hope. This suggests that
+although the build system is able to provide adequate parallelism, the
+underlying Linux OS and \verb|make| job scheduling fail to take
+advantage. To enable better scaling, potential sources of contention
+such as the filesystem must be examined.
 

### 2019-01-17 paper.tex
3.4

@@
 
-Google's protocol buffers project \cite{protobufs3.6.1} exhibited this behavior.  It's work was calculated
-to be about 16.7 minutes and it had a relatively
-long predicted span of 43.7 seconds, which was dominated by a call to g++ compiling
-a single file.  When the analysis applied in section \ref{sec:expensive} is applied
-to protocol buffers' build, we learn that the top three spans are dominated by the
-single file compilations of the files in fig. \ref{code:g++1}.  If the g++ compilations
-of these three files were not in the build's span the new span would be about 20s,
-roughly half of the original span.  Speeding up the compilation of these three files
-would be a good place to begin improving the parallelism of this build.
+Google's protocol buffers project \cite{protobufs3.6.1} exhibited this
+behavior.  It's work was calculated to be about 16.7 minutes and it
+had a relatively long predicted span of 43.7 seconds, which was
+dominated by a call to g++ compiling a single file.  When the analysis
+applied in section \ref{sec:expensive} is applied to protocol buffers'
+build, we learn that the top three spans are dominated by the single
+file compilations of the files in fig. \ref{code:g++1}.  If the g++
+compilations of these three files were not in the build's span the new
+span would be about 20s, roughly half of the original span.  Speeding
+up the compilation of these three files would be a good place to begin
+improving the parallelism of this build, perhaps by breaking
+individual files into multiple files. Our tool is automatically able
+to pinpoint where this would be profitable, and to show that further
+efforts past these three would have diminishing returns.
 
-\begin{figure}[H]
-\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
-  google/protobuf/descriptor.cc
-  google/protobuf/descriptor.pb.cc
-  google/protobuf/compiler/cpp/cpp\_message.cc
+\begin{figure}[t]
+\begin{Verbatim}[fontsize=\small]
+[span: 42.7; 38.7]
+ Longest part of span is `g++ descriptor.cc`
+Without this: [span: 30.45; 26.36]
+ Longest part of span is `g++ descriptor.pb.cc`
+Without this: [span: 28.73; 26.35]
+ Longest part of span is `g++ cpp_message.cc`
+Without this: [span: 19.78; 17.4]
 \end{Verbatim}
-\caption{The compilation of these 3 files dominate the longest 3 spans.}
+\caption{Output of our tool describing the protobuf build (simplified)}
 \label{code:g++1}

### 2019-01-17 paper.tex
Change subtitle.
spelling

@@
 to compile software projects faster. Fortunately, for most systems and
-languages, compiling software is a classic example of embarassing
+languages, compiling software is a classic example of embarrassing
 parallelism. As a result, developers who spend much of their time
@@
 \label{sec:analysis}
-In this section we describe the tools we use to analyse make builds, diagnose problems related
+In this section we describe the tools we use to analyze make builds, diagnose problems related
 to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
@@
 
-\subsection{Expensive command}
+\subsection{Too large command}
 \label{sec:expensive}
@@
 
-Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyse
+Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyze
 the span of zstd's build we saw that it is dominated by a single call to cc; see fig. \ref{code:cc1}.
@@
 \end{Verbatim}
-\caption{A long running cc command which was responsible for aproximately 100\% of zstd's span when the command in fig. \ref{code:cc1} is removed.
+\caption{A long running cc command which was responsible for approximately 100\% of zstd's span when the command in fig. \ref{code:cc1} is removed.
 This command compiled and linked 30 files.}
@@
 that potential parallelism in practice.  LLVM~\citep{LLVM:CGO04} is
-the largest project we analysed, with a calculated work of
+the largest project we analyzed, with a calculated work of
 approximately 12584 seconds, or 3.5 hours, and a predicted span of 305
@@
 advantage. To enable better scaling, potential sources of contention
-such as the filesystem must be examined.
+such as the file system must be examined.
 

### 2019-01-17 paper.tex
edits

@@
 
-In this section we discuss four problems with project builds that use make.  In section \ref{sec:expensive},
-we discuss
-builds whose spans are dominated by a single expensive command that prevents the build from
-achieving greater parallelism, look at a project that exhibits this behavior, and explain our
-our tool can diagnose this problem and identify areas for improvement.  In section \ref{sec:sequential}
-, we discuss what
-it means for a build to be unnecessarily sequential, and look at an example of a build that
-exhibits this behavior, and how our tool can diagnose this problem and identify areas for improvement.
-In section \ref{sec:scalability}, we discuss when a build should have ample available parallelism
-but fails to achieve that
-parallelism in practice, look at a project that exhibits this behavior, and discuss potential causes.
-In section \ref{sec:compiler} we discuss another instance of an expensive command dominating a
-build's span; in this case a
-slow compiler, we look at a project that exhibits this behavior and how our tool can diagnose it
-and identify areas for improvement.
+In this section we discuss four problems with project builds that use
+make.  In section \ref{sec:expensive}, we discuss builds whose spans
+are dominated by a single expensive command that prevents the build
+from achieving greater parallelism, look at a project that exhibits
+this behavior, and explain our our tool can diagnose this problem and
+identify areas for improvement.  In section \ref{sec:sequential} , we
+discuss what it means for a build to be unnecessarily sequential, and
+look at an example of a build that exhibits this behavior, and how our
+tool can diagnose this problem and identify areas for improvement.  In
+section \ref{sec:scalability}, we discuss when a build should have
+ample available parallelism but fails to achieve that parallelism in
+practice, look at a project that exhibits this behavior, and discuss
+potential causes.  In section \ref{sec:compiler} we discuss another
+instance of an expensive command dominating a build's span; in this
+case a slow compiler, we look at a project that exhibits this behavior
+and how our tool can automatically diagnose it and identify areas for improvement.
 
@@
 
-All project builds which lack the desired parallelism can be characterized as having too great of
-a span in relation to their work.  The span in a build system is the longest (in time) series of
-tasks that must be executed sequentially.  In some cases the span is not a series of similar length
-tasks, but is dominated by a single long running task.  In cases like these it is best to identify
-the most expensive task in the span, because reducing the running time of that task is the
-most likely way to make progress towards increasing the parallelism of the build.
+All project builds which lack the desired parallelism can be
+characterized as having too great of a span in relation to their work.
+The span in a build system is the longest (in time) series of tasks
+that must be executed sequentially.  In some cases the span is not a
+series of similar length tasks, but is dominated by a single long
+running task.  In cases like these it is best to identify the most
+expensive task in the span, because reducing the running time of that
+task is the most likely way to make progress towards increasing the
+parallelism of the build.
 
-To diagnose the problem of a span dominated by a single long running task, our tool can be used
-to learn the span path through the build graph and learn which target and shell commands
-are the most time intensive.  To see what sort of an improvement would occur in the span if the
-running time of the most expensive target in the span was to decrease, the tool will recalculate
-the span of the graph excluding its most expensive member any number of times.  This is useful
-in the case where decreasing the running time of a single task gives the build graph a new span
-path with a similarly long running time, which is dominated by a different long running command.
+To diagnose the problem of a span dominated by a single long running
+task, our tool can be used to learn the span path through the build
+graph and learn which target and shell commands are the most time
+intensive.  To see what sort of an improvement would occur in the span
+if the running time of the most expensive target in the span was to
+decrease, the tool will recalculate the span of the graph excluding
+its most expensive member any number of times.  This is useful in the
+case where decreasing the running time of a single task gives the
+build graph a new span path with a similarly long running time, but
+ dominated by a \emph{different} long running command.
 
-Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyze
-the span of zstd's build we saw that it is dominated by a single call to cc; see fig. \ref{code:cc1}.
+Facebook's zstd project~\cite{zstd1.3.7} has a build that exhibits
+this property.  When we used our tool to analyze the span of zstd's
+build we saw that it is dominated by a single call to \verb|cc|; see
+fig.~\ref{code:cc1}.
 
@@
 \end{Verbatim}
-\caption{A long running cc command which was responsible for about 97\% of zstd's span.  It compiled and linked 27 .c files.}
+\caption{This is responsible for about 97\% of zstd's span.  It compiles and links 27 .c files.}
 \label{code:cc1}
@@
 
-If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
-see fig. \ref{code:cc2}.
+If this command is removed from the build, the new span is similarly dominated by another single call to \verb|cc|; see fig. \ref{code:cc2}.
 \begin{figure}[H]
@@
 \end{Verbatim}
-\caption{A long running cc command which was responsible for approximately 100\% of zstd's span when the command in fig. \ref{code:cc1} is removed.
-This command compiled and linked 30 files.}
+\caption{A long running cc command responsible for approximately 100\%
+  of zstd's span when the command in fig. \ref{code:cc1} is removed.
+  This command compiles and links 30 files.}
 \label{code:cc2}
 \end{figure}
-If neither of these calls are members of the build's span, then our tool predicts the span would be
-approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
-This is about an 80\% reduction in the build's span, and the parallelism improves from 2.98 to 16.
+If neither of these calls are members of the build's span, then our
+tool predicts the span would be approximately 7.2 seconds rather than
+39 seconds, as with the original span, or 37 seconds, as with the
+second span.  This is an 80\% reduction in the build's span, and
+the parallelism improves from 2.98 to 16.
 
-To test this we manually replaced the cc command in fig. \ref{code:cc1} with the code in fig.
-\ref{code:cc3}.  The original command is replace by a new target for each object file and target
-that can build each object file in parallel and link them together afterwards.
+To test this we manually replaced the cc command in
+fig. \ref{code:cc1} with the code in fig.  \ref{code:cc3}.  The
+original command is replaced by a new target for each object file and
+target that can build each object file in parallel and link them
+together afterwards.
 
@@
 
-Similarly, the code in fig. \ref{code:cc2} is replaced by the code in fig. \ref{code:cc4}.  Once
-again each object file is explicitly declared as its own target and a target is declared that can
-build these object files in parallel and then link them together afterwards.
+Similarly, the code in fig. \ref{code:cc2} is replaced by the code in
+fig. \ref{code:cc4}.  Once again each object file is explicitly
+declared as its own target and a target is declared that can build
+these object files in parallel and then link them together afterwards.
 
@@
 
-Figure \ref{fig:zstd} shows that when the modified build was run with 1 to 32 jobs,
-the predicted speedup was observed.  An additional performance improvement that was not predicted
-is that the work of the new build decreased from approximately 120 seconds to approximately 80 seconds.
-This is because each of the object files is declared as its own target, so make can consider whether
-or not the file already exists and if it is necessary to rebuild it.  In this case, it turns out that
-the object files being built as part of the cc command in fig. \ref{code:cc1} had already been built
-by make; making it unnecessary to build them again.  This cuts the effect time of the code in fig.
-\ref{code:cc1} from about 40 seconds to about 2 seconds; just the time to link the object files.
+Figure \ref{fig:zstd} shows that when the modified build was run with
+1 to 32 jobs, the predicted speedup was observed.  An additional
+performance improvement that was not predicted is that the work of the
+new build decreased from approximately 120 seconds to approximately 80
+seconds.  This is because each of the object files is declared as its
+own target, so \verb|make| can consider whether or not the file
+already exists and if it is necessary to rebuild it.  In this case, it
+turns out that the object files being built as part of the cc command
+in fig. \ref{code:cc1} had already been built; making it unnecessary
+to build them again.  This cuts the effective time of the code in fig.
+\ref{code:cc1} from 40 seconds to 2 seconds---just the time to link
+the object files.
 
-\subsection{Unnecessarily sequential}
+\subsection{Unnecessarily sequential rules}
 \label{sec:sequential}
